On the uniqueness and stability of dictionaries for sparse representation of noisy signals

### 2017-03-29 manuscripts/PNAS/chazthm_pnas.tex
more good ol fashioned SR-ing

More formally, let $\mathbf{A} \in \mathbb R^{n \times m}$ be a matrix with columns $\mathbf{A}_j$ ($j = 1,\ldots,m$) and let dataset $Z$ consist of measurements:
\begin{align}\label{LinearModel}
\mathbf{z}_i = \mathbf{A}\mathbf{x}_i + \mathbf{n}_i,\indent \text{$i=1,\ldots,N$},
\end{align}
for $k$-\emph{sparse} $\mathbf{x}_i \in \mathbb{R}^m$ having at most $k$ nonzero entries and \emph{noise} $\mathbf{n}_i \in \mathbb{R}^n$, with bounded norm $\| \mathbf{n}_i \|_2 \leq  \eta$ representing our combined worst-case uncertainty in  measuring $\mathbf{A}\mathbf{x}_i$.
The precise mathematical problem addressed here is the following.
% We address the uniqueness properties for the following linear coding question in signal processing.

@@

%  We introduce the following terminology to handle $\eta > 0$. 
\begin{definition}\label{maindef}
Fix $Y = \{ \mathbf{y}_1, \ldots, \mathbf{y}_N\} \subset \mathbb{R}^n$. We say $Y$ has a \textbf{$k$-sparse linear representation in $m$ dimensions} if there exists a matrix $\mathbf{A}$ and $k$-sparse $\mathbf{x}_1, \ldots, \mathbf{x}_N \in \mathbb{R}^m$ such that $\mathbf{y}_i = \mathbf{A}\mathbf{x}_i$ for all $i$. %Equivalently, $\mathbf{Y} = \mathbf{A}\mathbf{X}$ for $\mathbf{Y}$ and $\mathbf{X}$ having the $\mathbf{y}_i$ and $\mathbf{x}_i$ as their columns, respectively. 
This representation is \textbf{stable} if for every $\delta_1, \delta_2 \geq 0$, there exists $\varepsilon(\delta_1, \delta_2) \geq 0$ (with $\varepsilon > 0$ when  $\delta_1, \delta_2 > 0$) such that if $\mathbf{B}$ and $k$-sparse $\mathbf{\hat x}_1, \ldots, \mathbf{\hat x}_N \in \mathbb{R}^m$ satisfy:
\begin{align*}
\|\mathbf{A}\mathbf{x}_i - \mathbf{B}\mathbf{\hat x}_i\|_2 \leq \varepsilon, \ \ \text{for all $i$},
\end{align*}
%
then there is some permutation matrix $\mathbf{P}$ and invertible diagonal matrix $\mathbf{D}$ such that for all $i, j$:
\begin{align}\label{def1}
\|\mathbf{A}_j - (\mathbf{BPD})_j\|_2 \leq \delta_1 \ \ \text{and} \ \ \|\mathbf{x}_i - \mathbf{D}^{-1}\mathbf{P}^{\top}\mathbf{\hat x}_i\|_1 \leq \delta_2.
\end{align}
\end{definition}
% Kato mentions a distance between n-tuples that minimizes the maximum difference over all possible orderings of one of the n-tuples.
% Make just 'sparse representation in m-dimensions? (i.e. with k unspecified?)

@@

\begin{proposition}
For every $k < m$, there is a regular $k$-uniform hypergraph that satisfies the singleton intersection property.
\end{proposition}
Consider, for instance, the set of consecutive intervals of length $k$ in some cyclic order on $[m]$.

We say that a set of $k$-sparse vectors is in \emph{general linear position} when any $k$ of them are linearly independent.
The following is the technical statement of our main result.

\begin{theorem}\label{DeterministicUniquenessTheorem}
% Fix positive integers $k, \ell, m, m', n, p$ satisfying $k < m$ and $p \leq m \leq m' \leq (\ell m - p) / (\ell - 1)$. 
Fix integers $n, \ell$, and $k < m \leq \hat m$. Suppose $\mathbf{A}~\in~\mathbb{R}^{n \times m}$ satisfies the spark condition \eqref{SparkCondition} and that $k$-sparse \mbox{$\mathbf{x}_1, \ldots, \mathbf{x}_N \in \mathbb{R}^m$} contain at least \mbox{$(k-1){\hat m \choose k}+1$} vectors supported in general linear position on each set in some $\ell$-regular $E \subseteq {[m] \choose k}$ satisfying the SIP.  There exists a constant $C_1 > 0$ for which the following holds for all $\varepsilon < L_2(\mathbf{A}) / C_1$.

Every matrix $\mathbf{B} \in \mathbb{R}^{n \times \hat m}$ and $k$-sparse $\mathbf{\hat x}_1, \ldots, \mathbf{\hat x}_N \in \mathbb{R}^{\hat m}$ with  \mbox{$\|\mathbf{A}\mathbf{x}_i - \mathbf{B}\mathbf{\hat x}_i\|_2 \leq \varepsilon$} for all $i$ necessarily satisfies:
\begin{align}\label{Cstable}
\|\mathbf{A}_{\phi(j)} - (\mathbf{B}_{J}\mathbf{PD})_j\|_2 \leq C_1 \varepsilon, \ \ \text{for $j = 1, \ldots, |J|$},
\end{align}
%
for some $J \subseteq [\hat m]$ of size \mbox{$\lfloor \hat m - \ell(\hat m - m) \rfloor$}, injective map $\phi$, permutation matrix $\mathbf{P}$, and invertible diagonal matrix $\mathbf{D}$.
%\footnote{We note that the condition $\varepsilon < L_2(\mathbf{A})$ above is necessary. When \mbox{$\mathbf{A}$ = $I$} and $\mathbf{x}_i = \mathbf{e}_i$, it turns out that $C_1 = 1$ and there is a $\mathbf{B}$ and $1$-sparse $\mathbf{\hat x}_i$ with $\|\mathbf{A}\mathbf{x}_i - \mathbf{B}\mathbf{\hat x}_i\|_2 \leq \varepsilon$ violating \eqref{Cstable}.}

Furthermore, if $\varepsilon < L_{2k}(\mathbf{A}) / C_1$, then $\mathbf{B}_J$ also satisfies \eqref{SparkCondition} with $L_{2k}(\mathbf{B}_J\mathbf{PD}) \geq L_{2k}(\mathbf{A}) - C_1 \varepsilon$, and:
\begin{align}\label{b-PDa}
\|\mathbf{x}_i - \mathbf{D}^{-1}\mathbf{P}^{\top}\mathbf{\hat x}_i\|_1 &\leq  \left( \frac{ 1+C_1 \|\mathbf{x}_i\|_1 }{ L_{2k}(\mathbf{A}) -  C_1\varepsilon } \right) \varepsilon,
 \ \ \text{for all $i \in [N]$}.
\end{align}
\end{theorem}

In other words, the smaller the difference $\hat m - m$, the more columns and coefficients of the original $n \times m$ dictionary $\mathbf{A}$ and $m$-dimensional codes $\mathbf{x}_i$ contained (up to noise) in the appropriately scaled $n \times \hat m$ dictionary $\mathbf{B}$ and $\hat m$-dimensional codes $\mathbf{\hat x}_i$. In particular, when $\hat m = m$ all columns of $\mathbf{A}$ and coefficients of each $\mathbf{x}_i$ recoverable, as per Def.~\ref{maindef}. For expositional clarity, we delay defining the (explicit) constant $C_1 = C_1(\mathbf{A}, \{\mathbf{x}_i\}_{i=1}^N, E)$ until Section \ref{DUT} (\eqref{Cdef}).  % , since it requires additional definitions not critical to the statement of our results.

@@

\begin{corollary}\label{DeterministicUniquenessCorollary}
Given $n, m$, $k < m$, and a regular hypergraph $E \subseteq {[m] \choose k}$ of rank $k$ with the SIP, there are $N =  |E| \left[ (k-1){m \choose k}+1 \right]$ vectors \mbox{$\mathbf{x}_1, \ldots, \mathbf{x}_N \in \mathbb{R}^m$} such that every matrix $\mathbf{A} \in \mathbb{R}^{n \times m}$ satisfying \eqref{SparkCondition} generates a dataset $Y = \{\mathbf{A}\mathbf{x}_1, \ldots, \mathbf{A}\mathbf{x}_N\}$ with a stable $k$-sparse representation in $m$ dimensions.
\end{corollary}
\begin{proof}
Something quick here.
\end{proof}

%\begin{proposition}
%If $E$ is an $\ell$-regular hypergraph of rank $k$ satisfying the singleton intersection property then $|E| \geq \ell m/k$. 
%\end{proposition}
Note that if $E$ is an $\ell$-regular hypergraph of rank $k$ satisfying the singleton intersection property then $|E| \geq \ell m/k$.
This is easy to see by noting that $|E|k \geq \sum_{S \in E}|S| = \sum_{i \in [m]} \deg(i) = \ell m$. In certain cases, there exist simple constructions achieving this minimum; for example, when $k = \sqrt{m}$ one can take $E$ to be the rows and columns formed by arranging the elements of $[m]$ in a square grid.

It is straightforward to provide a probabilistic extension of Thm.~\ref{DeterministicUniquenessTheorem} using the following fact in random matrix theory.  A random matrix $\mathbf{A} \in \mathbb{R}^{n \times m}$ satisfies \eqref{SparkCondition} with probability one (or with ``high probability'' for discrete variables) 
provided:
\begin{align}\label{CScondition}
n \geq \gamma k\log\left(\frac{m}{k}\right),
\end{align}
%
in which $\gamma$ is a positive constant that depends on the particular distribution from which the entries of $\mathbf{A}$ are sampled i.i.d. (many ensembles suffice, e.g. \cite[Sec.~4]{Baraniuk08}).

We motivate our next theorem by the following observation:  the spark condition can be made explicit.  Let $\mathbf{A}$  be the $n \times m$ matrix of $nm$ indeterminates $A_{ij}$. When real numbers are substituted for $A_{ij}$, the resulting matrix satisfies \eqref{SparkCondition} if and only if the following polynomial is nonzero:
\begin{align*}
f(\mathbf{A}) := \prod_{S \in {[m] \choose k}} \sum_{S' \in {[n] \choose k}} (\det \mathbf{A}_{S',S})^2,
\end{align*}
%
where for any $S' \in {[n] \choose k}$ and $S \in {[m] \choose k}$, the symbol $\mathbf{A}_{S',S}$ denotes the submatrix of entries $A_{ij}$ with $(i,j) \in S' \times S$.   We note that the large number of terms in this product is likely necessary due to the NP-hardness of deciding whether a given matrix $A$ satisfies the spark condition \cite{tillmann2014computational}.

@@

\begin{theorem}\label{robustPolythm} %label is fucking up formatting..???
Fix $n, m$, and $k < m$. There is a polynomial in the entries of $\mathbf{A}$ and the $\mathbf{x}_i$ with the following property:  if the polynomial evaluates to a nonzero number and at least \mbox{$(k-1){m \choose k}+1$} of the resulting vectors $\mathbf{x}_i$ are supported on each $S \in E$ for some regular $k$-uniform $E \subseteq 2^{[m]}$ satisfying the singleton intersection property, then $Y$ has a stable $k$-sparse representation in $\mathbb{R}^m$ (Def.~\ref{Uniqueness}). In particular, either no substitutions impart to $Y$ this property or all but a Borel set of measure zero do. 
\end{theorem}

@@

\begin{corollary}\label{ProbabilisticCor}
Fix $n, m$, and $k$ satisfying \eqref{CScondition} for $\gamma = \gamma_0$, and let the entries of the matrix $\mathbf{A} \in \mathbb{R}^{n \times m}$ and $k$-sparse vectors $\mathbf{x}_1, \ldots, \mathbf{x}_N \in \mathbb{R}^m$ be drawn independently from probability measures absolutely continuous with respect to the standard Borel measure $\mu$. If at least $(k-1){m \choose k} + 1$ of the vectors $\mathbf{x}_i$ are supported on each $S \in E$ for some regular $k$-uniform $E \subseteq 2^{[m]}$ satisfying the singleton intersection property, then $Y$ has a stable $k$-sparse representation in $\mathbb{R}^m$ with probability one.
\end{corollary}

@@

\begin{proof}[Proof ($k=1$)]
The only 1-uniform hypergraph satisfying the singleton intersection property is $[m]$; hence, we have $\mathbf{x}_i = c_i \mathbf{e}_i$ for $c_i \in \mathbb{R} \setminus \{\mathbf{0}\}$, $i \in [m]$. In this case we may take $C_1 = 1/ \min_{\ell \in [m]} |c_{\ell}|$.

Fix $\mathbf{A} \in \mathbb{R}^{n \times m}$ satisfying \eqref{SparkCondition} and suppose that for some $\mathbf{B} \in \mathbb{R}^{n \times \hat m}$ ($\hat m \geq m$) and $1$-sparse $\mathbf{\hat x}_i \in \mathbb{R}^{\hat m}$ we have  $\|\mathbf{A}\mathbf{x}_i - \mathbf{B}\mathbf{\hat x}_i\|_2 \leq \varepsilon < L_2(\mathbf{A}) / C_1$ for all $i$. Then there are $c'_1, \ldots, c'_m \in \mathbb{R}$ and $\pi: [m] \to [\hat m]$ with:
\begin{align}\label{1D}
\|c_j\mathbf{A}_j - c'_j\mathbf{B}_{\pi(j)}\|_2 \leq \varepsilon, \ \ \text{for all $j$}.
\end{align} 
Note that $c'_j \neq 0$ for all $j$ since otherwise by \eqref{delrho} and definition of $C_1$ we would have $\|c_j\mathbf{A}_j\|_2 < \min_{\ell \in [m]}\|c_{\ell}\mathbf{A}_{\ell}\|_2$.

@@

We note the following facts about $d$. For subspaces $U,V \subseteq \mathbb{R}^m$, we have \cite[Cor.~2.6]{Kato2013}:
\begin{align}\label{dimLem}
d(U,V) < 1 \implies \dim(U) \leq \dim(V).
\end{align}
%
and \cite[Lem.~3.2]{Morris10}:
\begin{align}\label{eqdim}
\dim(U) = \dim(V) \implies d(U,V) = d(V,U).
\end{align}

@@

\begin{lemma}\label{MainLemma}
Fix integers $n$ and $k < m$, and suppose $\mathbf{A} \in \mathbb{R}^{n \times m}$ satisfies the spark condition \eqref{SparkCondition}. There exists a constant $C_2 > 0$ for which the following holds for all $\varepsilon < L_2(\mathbf{A}) / C_2$. If for some  $\mathbf{B} \in \mathbb{R}^{n \times m}$ and regular $E \subseteq2^{[m]}$ satisfying the singleton intersection property there exists a size-preserving map $\pi: E \mapsto 2^{[m]}$ satisfying:
\begin{align}\label{GapUpperBound}
d(\text{\rm Span}\{\mathbf{A}_{S}\}, \text{\rm Span}\{\mathbf{B}_{\pi(S)}\}) \leq \varepsilon, \ \ \text{for all $S \in E$},
\end{align}
%
then there exists some permutation matrix $\mathbf{P}$ and invertible diagonal matrix $\mathbf{D}$ such that:
\begin{align}\label{MainLemmaBPD}
\|\mathbf{A}_j - (\mathbf{BPD})_j\|_2 \leq C_2 \varepsilon, \indent \text{for } j \in [m].
\end{align}
\end{lemma}

>>>>

More formally, let $\mathbf{A} \in \mathbb R^{n \times m}$ be a matrix with columns $\mathbf{A}_j$ ($j = 1,\ldots,m$) and let dataset $Z$ consist of measurements:
\begin{align}\label{LinearModel}
\mathbf{z}_i = \mathbf{A}\mathbf{x}_i + \mathbf{n}_i,\ \ \  \text{$i=1,\ldots,N$},
\end{align}
for $k$-\emph{sparse} $\mathbf{x}_i \in \mathbb{R}^m$ having at most $k$ nonzero entries and \emph{noise} $\mathbf{n}_i \in \mathbb{R}^n$, with bounded norm $\| \mathbf{n}_i \|_2 \leq  \eta$ representing our combined worst-case uncertainty in  measuring $\mathbf{A}\mathbf{x}_i$.
The precise mathematical problem addressed here is the following.
% We address the uniqueness properties for the following linear coding question in signal processing.

@@

%  We introduce the following terminology to handle $\eta > 0$. 
\begin{definition}\label{maindef}
Fix $Y = \{ \mathbf{y}_1, \ldots, \mathbf{y}_N\} \subset \mathbb{R}^n$. We say $Y$ has a \textbf{$k$-sparse linear representation in $m$ dimensions} if there exists a matrix $\mathbf{A}$ and $k$-sparse $\mathbf{x}_1, \ldots, \mathbf{x}_N \in \mathbb{R}^m$ such that $\mathbf{y}_i = \mathbf{A}\mathbf{x}_i$ for all $i$. %Equivalently, $\mathbf{Y} = \mathbf{A}\mathbf{X}$ for $\mathbf{Y}$ and $\mathbf{X}$ having the $\mathbf{y}_i$ and $\mathbf{x}_i$ as their columns, respectively. 
This representation is \textbf{stable} if for every $\delta_1, \delta_2 \geq 0$, there exists $\varepsilon(\delta_1, \delta_2) \geq 0$ (with $\varepsilon > 0$ when  $\delta_1, \delta_2 > 0$) such that if $\mathbf{B}$ and $k$-sparse $\mathbf{\hat x}_1, \ldots, \mathbf{\hat x}_N \in \mathbb{R}^m$ satisfy:
\begin{align*}
\|\mathbf{A}\mathbf{x}_i - \mathbf{B}\mathbf{\hat x}_i\|_2 \leq \varepsilon,\ \ \   \text{for all $i$},
\end{align*}
%
then there is some permutation matrix $\mathbf{P}$ and invertible diagonal matrix $\mathbf{D}$ such that for all $i, j$:
\begin{align}\label{def1}
\|\mathbf{A}_j - (\mathbf{BPD})_j\|_2 \leq \delta_1 \ \ \text{and} \ \ \|\mathbf{x}_i - \mathbf{D}^{-1}\mathbf{P}^{\top}\mathbf{\hat x}_i\|_1 \leq \delta_2.
\end{align}
\end{definition}
% Kato mentions a distance between n-tuples that minimizes the maximum difference over all possible orderings of one of the n-tuples.
% Make just 'sparse representation in m-dimensions? (i.e. with k unspecified?)

@@

It is easy to verify that for every $k < m$, there is a regular $k$-uniform hypergraph that satisfies the SIP; for instance, consecutive intervals of length $k$ in some cyclic order on $[m]$.
%\begin{proposition}
%For every $k < m$, there is a regular $k$-uniform hypergraph that satisfies the singleton intersection property.
%\end{proposition}

We say that a set of $k$-sparse vectors is in \emph{general linear position} when any $k$ of them are linearly independent.
The following is the precise statement of our main result.

\begin{theorem}\label{DeterministicUniquenessTheorem}
Fix integers $n, \ell$, and $k < m$. Suppose $\mathbf{A}~\in~\mathbb{R}^{n \times m}$ satisfies the spark condition \eqref{SparkCondition} and that $k$-sparse \mbox{$\mathbf{x}_1, \ldots, \mathbf{x}_N \in \mathbb{R}^m$} contain at least \mbox{$(k-1){\hat m \choose k}+1$} vectors supported in general linear position on each set in some $\ell$-regular $E \subseteq {[m] \choose k}$ with the SIP.  There exists a constant $C_1 > 0$ for which the following holds for all $\varepsilon < L_2(\mathbf{A}) / C_1$.

Every matrix $\mathbf{B} \in \mathbb{R}^{n \times m}$ and $k$-sparse $\mathbf{\hat x}_1, \ldots, \mathbf{\hat x}_N \in \mathbb{R}^{m}$ with  \mbox{$\|\mathbf{A}\mathbf{x}_i - \mathbf{B}\mathbf{\hat x}_i\|_2 \leq \varepsilon$} for all $i$ necessarily satisfies:
\begin{align}\label{Cstable}
\|\mathbf{A}_j - (\mathbf{B}\mathbf{PD})_j\|_2 \leq C_1 \varepsilon,\ \ \   \text{for all $j \in [m]$},
\end{align}
for some permutation matrix $\mathbf{P}$ and invertible diagonal $\mathbf{D}$.

Furthermore, if $\varepsilon < L_{2k}(\mathbf{A}) / C_1$, then $\mathbf{B}$ also satisfies \eqref{SparkCondition} with $L_{2k}(\mathbf{B}\mathbf{PD}) \geq L_{2k}(\mathbf{A}) - C_1 \varepsilon$, and:
\begin{align}\label{b-PDa}
\|\mathbf{x}_i - \mathbf{D}^{-1}\mathbf{P}^{\top}\mathbf{\hat x}_i\|_1 &\leq  \left( \frac{ 1+C_1 \|\mathbf{x}_i\|_1 }{ L_{2k}(\mathbf{A}) -  C_1\varepsilon } \right) \varepsilon,\ \ \   \text{for all $i \in [N]$}.
\end{align}
\end{theorem}

%\begin{theorem}\label{DeterministicUniquenessTheorem}
%Fix integers $n, \ell$, and $k < m \leq \hat m$. Suppose $\mathbf{A}~\in~\mathbb{R}^{n \times m}$ satisfies the spark condition \eqref{SparkCondition} and that $k$-sparse \mbox{$\mathbf{x}_1, \ldots, \mathbf{x}_N \in \mathbb{R}^m$} contain at least \mbox{$(k-1){\hat m \choose k}+1$} vectors supported in general linear position on each set in some $\ell$-regular $E \subseteq {[m] \choose k}$ satisfying the SIP.  There exists a constant $C_1 > 0$ for which the following holds for all $\varepsilon < L_2(\mathbf{A}) / C_1$.
%
%Every matrix $\mathbf{B} \in \mathbb{R}^{n \times \hat m}$ and $k$-sparse $\mathbf{\hat x}_1, \ldots, \mathbf{\hat x}_N \in \mathbb{R}^{\hat m}$ with  \mbox{$\|\mathbf{A}\mathbf{x}_i - \mathbf{B}\mathbf{\hat x}_i\|_2 \leq \varepsilon$} for all $i$ necessarily satisfies:
%\begin{align}\label{Cstable}
%\|\mathbf{A}_{\phi(j)} - (\mathbf{B}_{J}\mathbf{PD})_j\|_2 \leq C_1 \varepsilon,\ \ \   \text{for $j = 1, \ldots, |J|$},
%\end{align}
%%
%for some $J \subseteq [\hat m]$ of size \mbox{$\lfloor \hat m - \ell(\hat m - m) \rfloor$}, injective map $\phi$, permutation matrix $\mathbf{P}$, and invertible diagonal matrix $\mathbf{D}$.
%%\footnote{We note that the condition $\varepsilon < L_2(\mathbf{A})$ above is necessary. When \mbox{$\mathbf{A}$ = $I$} and $\mathbf{x}_i = \mathbf{e}_i$, it turns out that $C_1 = 1$ and there is a $\mathbf{B}$ and $1$-sparse $\mathbf{\hat x}_i$ with $\|\mathbf{A}\mathbf{x}_i - \mathbf{B}\mathbf{\hat x}_i\|_2 \leq \varepsilon$ violating \eqref{Cstable}.}
%
%Furthermore, if $\varepsilon < L_{2k}(\mathbf{A}) / C_1$, then $\mathbf{B}_J$ also satisfies \eqref{SparkCondition} with $L_{2k}(\mathbf{B}_J\mathbf{PD}) \geq L_{2k}(\mathbf{A}) - C_1 \varepsilon$, and:
%\begin{align}\label{b-PDa}
%\|\mathbf{x}_i - \mathbf{D}^{-1}\mathbf{P}^{\top}\mathbf{\hat x}_i\|_1 &\leq  \left( \frac{ 1+C_1 \|\mathbf{x}_i\|_1 }{ L_{2k}(\mathbf{A}) -  C_1\varepsilon } \right) \varepsilon,\ \ \   \text{for all $i \in [N]$}.
%\end{align}
%\end{theorem}

%In other words, the smaller the difference $\hat m - m$, the more columns and coefficients of the original $n \times m$ dictionary $\mathbf{A}$ and $m$-dimensional codes $\mathbf{x}_i$ contained (up to noise) in the appropriately scaled $n \times \hat m$ dictionary $\mathbf{B}$ and $\hat m$-dimensional codes $\mathbf{\hat x}_i$. In particular, when $\hat m = m$ all columns of $\mathbf{A}$ and coefficients of each $\mathbf{x}_i$ recoverable, as per Def.~\ref{maindef}.

For expositional clarity, we delay defining the (explicit) constant $C_1 = C_1(\mathbf{A}, \{\mathbf{x}_i\}_{i=1}^N, E)$ until Section \ref{DUT} (\eqref{Cdef1}).  % , since it requires additional definitions not critical to the statement of our results.

@@

\begin{corollary}\label{DeterministicUniquenessCorollary}
Given $n, m$, $k < m$, and a regular hypergraph $E \subseteq {[m] \choose k}$ with the SIP, there are $N =  |E| \left[ 1 + (k-1){m \choose k} \right]$ vectors \mbox{$\mathbf{x}_1, \ldots, \mathbf{x}_N \in \mathbb{R}^m$} such that every matrix $\mathbf{A} \in \mathbb{R}^{n \times m}$ satisfying \eqref{SparkCondition} generates a dataset $Y = \{\mathbf{A}\mathbf{x}_1, \ldots, \mathbf{A}\mathbf{x}_N\}$ with a stable $k$-sparse representation in $m$ dimensions.
\end{corollary}
\begin{proof}
Something quick here.
\end{proof}

%\begin{proposition}
%If $E$ is an $\ell$-regular hypergraph of rank $k$ satisfying the singleton intersection property then $|E| \geq \ell m/k$. 
%\end{proposition}
Note that if $E$ is an $\ell$-regular hypergraph of rank $k$ satisfying the singleton intersection property, then $|E| \geq \ell m/k$.
This is easy to see from $|E|k \geq \sum_{S \in E}|S| = \sum_{i \in [m]} \deg(i) = \ell m$. In certain cases, there exist simple constructions achieving this minimum; for example, when $k = \sqrt{m}$ one can take $E$ to be the rows and columns formed by arranging the elements of $[m]$ in a square grid.  In light of Cor.~\ref{DeterministicUniquenessCorollary}, it would be  interesting problem to determine minimal such $E$.

There are other less direct consequences of Thm.~\ref{DeterministicUniquenessTheorem} to stability in sparse linear coding.  The following implication relates the result to more common optimization formulations of the problem.

\begin{theorem}
Fix integers $n, \ell$ and $k < m$. Suppose $\mathbf{A}~\in~\mathbb{R}^{n \times m}$ satisfies the spark condition and that $k$-sparse \mbox{$\mathbf{x}_1, \ldots, \mathbf{x}_N \in \mathbb{R}^m$} contain at least \mbox{$(k+1)\left[ 1 + (k-1){ m \choose k} \right]$} vectors supported in general linear position on each set in some $\ell$-regular $E \subseteq {[m] \choose k}$ satisfying the SIP. There exists a constant $C_1 > 0$ for which the following holds for all $\varepsilon < L_2(\mathbf{A}) / C_1$.

Every solution to the constrained $\ell_0$ optimization problem:
\begin{align}\label{minsum}
\arg \min_{ \substack{ \mathbf{B} \in \mathbb{R}^{n \times m} \\ \mathbf{\hat x}_i, \ldots, \mathbf{\hat x}_N \in \mathbb{R}^{m} } } \sum_{i = 1}^N \|\mathbf{\hat x}_i\|_0 \ \ \text{s.t.} \ \ \|\mathbf{A}\mathbf{x}_i - \mathbf{B}\mathbf{\hat x}_i\|_2 \leq \varepsilon, \ \ \text{$i \in [N]$},
\end{align}
necessarily satisfies \eqref{Cstable}.
\end{theorem}

\begin{proof}
[*** Incomplete ***]
We argue for a lower bound on the number of $k$-sparse vectors $\mathbf{\hat x}_i$ and then apply Thm.~\ref{DeterministicUniquenessCorollary}. Let $n_p$ be the number of vectors $\mathbf{\hat x}_i$ satisfying $\|\mathbf{\hat x}_i\|_0 = p$, so that $\sum_{p = 1}^{\hat m} n_p = N$. Since the $\mathbf{x}_i$ are all $k$-sparse, by \eqref{minsum} we have:
\begin{align}
k \sum_{p = 1}^{m} n_p = kN \geq \sum_{i=1}^N \|\mathbf{x}_i\|_0 \geq \sum_{i=1}^N \|\mathbf{\hat x}_i\|_0 = \sum_{p=1}^{m} p n_p 
\end{align}
\[\implies \sum_{p \leq k} (k-p)n_p \geq \sum_{p > k} (p-k) n_p.\]
Hence,
\begin{align}
k \sum_{p \leq k} n_p \geq \sum_{p \leq k} (k-p)n_p \geq \sum_{p > k} (p-k) n_p \geq \sum_{p > k} n_p
\end{align}
\[ = N - \sum_{p \leq k} n_p \implies \sum_{p \leq k} n_p \geq N / (k+1).\]

That is, at least $N / (k+1)$ of the vectors $\mathbf{\hat x}_i$ are $k$-sparse. Therefore, there is some $N' \subseteq [N]$ for which for each $S \in E$ there are $(k-1){m \choose k}+1$ of the vectors $\mathbf{x}_i$ with $i \in [N']$ supported on $S$ and for which $\mathbf{\hat x}_i$ is $k$-sparse for all $i \in [N']$. By Thm.~\ref{DeterministicUniquenessCorollary}, \eqref{Cstable} follows.
\end{proof}

Another straightforward application of Thm.~\ref{DeterministicUniquenessTheorem} is a probabilistic extension that uses the following fact in random matrix theory.  A random matrix $\mathbf{A} \in \mathbb{R}^{n \times m}$ satisfies \eqref{SparkCondition} with probability one (or with ``high probability'' for discrete variables) 
provided:
\begin{align}\label{CScondition}
n \geq \gamma k\log\left(\frac{m}{k}\right),
\end{align}
in which $\gamma$ is a positive constant that depends on the particular distribution from which the entries of $\mathbf{A}$ are sampled i.i.d. (many ensembles suffice, e.g. \cite[Sec.~4]{Baraniuk08}).

In fact, 
% We motivate our next theorem by the following observation:  
the spark condition can be made explicit.  Let $\mathbf{A}$  be the $n \times m$ matrix of $nm$ indeterminates $A_{ij}$. When real numbers are substituted for $A_{ij}$, the resulting matrix satisfies \eqref{SparkCondition} if and only if the following polynomial is nonzero:
\begin{align*}
f(\mathbf{A}) := \prod_{S \in {[m] \choose k}} \sum_{S' \in {[n] \choose k}} (\det \mathbf{A}_{S',S})^2,
\end{align*}
%
where for any $S' \in {[n] \choose k}$ and $S \in {[m] \choose k}$, the symbol $\mathbf{A}_{S',S}$ denotes the submatrix of entries $A_{ij}$ with $(i,j) \in S' \times S$.   We note that the large number of terms in this product is likely necessary due to the NP-hardness of deciding whether a given matrix $A$ satisfies the spark condition \cite{tillmann2014computational}.

@@

\begin{theorem}\label{robustPolythm} %label is fucking up formatting..???
Fix $n, m$, and $k < m$. There is a polynomial in the entries of $\mathbf{A}$ and the $\mathbf{x}_i$ with the following property:  if the polynomial evaluates to a nonzero number and at least \mbox{$(k-1){m \choose k}+1$} of the resulting vectors $\mathbf{x}_i$ are supported on each $S \in E$ for some regular $E \subseteq {[m] \choose k}$ with the SIP, then $Y$ has a stable $k$-sparse representation in $\mathbb{R}^m$ (Def.~\ref{maindef}). In particular, either no substitutions impart to $Y$ this property or all but a Borel set of measure zero do. 
\end{theorem}

@@

\begin{corollary}\label{ProbabilisticCor}
Fix $n, m$, and $k$ satisfying \eqref{CScondition} for $\gamma = \gamma_0$, and let the entries of $\mathbf{A} \in \mathbb{R}^{n \times m}$ and $k$-sparse $\mathbf{x}_1, \ldots, \mathbf{x}_N \in \mathbb{R}^m$ be drawn independently from probability measures absolutely continuous with respect to the standard Borel measure $\mu$. If at least $(k-1){m \choose k} + 1$ of the vectors $\mathbf{x}_i$ are supported on each $S \in E$ for a regular $E \subseteq {[m] \choose k}$ with the SIP, then $Y$ has a stable $k$-sparse representation in $\mathbb{R}^m$ with probability one.
\end{corollary}

@@

\begin{proof}[Proof of Thm.~\ref{DeterministicUniquenessCorollary} for $k=1$]
The only 1-uniform hypergraph with the SIP is $[m]$; hence, we have $\mathbf{x}_i = c_i \mathbf{e}_i$ for $c_i \in \mathbb{R} \setminus \{\mathbf{0}\}$, $i \in [m]$. In this case, we may take $C_1 = 1/ \min_{\ell \in [m]} |c_{\ell}|$.

Fix $\mathbf{A} \in \mathbb{R}^{n \times m}$ satisfying \eqref{SparkCondition} and suppose that for some $\mathbf{B} \in \mathbb{R}^{n \times \hat m}$ ($\hat m \geq m$) and $1$-sparse $\mathbf{\hat x}_i \in \mathbb{R}^{\hat m}$ we have  $\|\mathbf{A}\mathbf{x}_i - \mathbf{B}\mathbf{\hat x}_i\|_2 \leq \varepsilon < L_2(\mathbf{A}) / C_1$ for all $i$. Then there are $c'_1, \ldots, c'_m \in \mathbb{R}$ and $\pi: [m] \to [\hat m]$ with:
\begin{align}\label{1D}
\|c_j\mathbf{A}_j - c'_j\mathbf{B}_{\pi(j)}\|_2 \leq \varepsilon,\ \ \   \text{for all $j$}.
\end{align} 
Note that $c'_j \neq 0$ for all $j$ since otherwise by \eqref{delrho} and definition of $C_1$ we would have $\|c_j\mathbf{A}_j\|_2 < \min_{\ell \in [m]}\|c_{\ell}\mathbf{A}_{\ell}\|_2$.

@@

We note the following facts about $d$. For subspaces $U,V \subseteq \mathbb{R}^m$, we have \cite[Cor.~2.6]{Kato2013}:
\begin{align}\label{dimLem}
d(U,V) < 1 \implies \dim(U) \leq \dim(V),
\end{align}
%
and \cite[Lem.~3.2]{Morris10}:
\begin{align}\label{eqdim}
\dim(U) = \dim(V) \implies d(U,V) = d(V,U).
\end{align}

@@

\begin{lemma}\label{MainLemma}
Fix integers $n$ and $k < m$, and suppose $\mathbf{A} \in \mathbb{R}^{n \times m}$ satisfies the spark condition \eqref{SparkCondition}. There exists a constant $C_2 > 0$ for which the following holds for all $\varepsilon < L_2(\mathbf{A}) / C_2$. If for some  $\mathbf{B} \in \mathbb{R}^{n \times m}$ and regular $E \subseteq2^{[m]}$ satisfying the singleton intersection property there exists a size-preserving map $\pi: E \mapsto 2^{[m]}$ satisfying:
\begin{align}\label{GapUpperBound}
d(\text{\rm Span}\{\mathbf{A}_{S}\}, \text{\rm Span}\{\mathbf{B}_{\pi(S)}\}) \leq \varepsilon,\ \ \   \text{for all $S \in E$},
\end{align}
%
then there exists some permutation matrix $\mathbf{P}$ and invertible diagonal matrix $\mathbf{D}$ such that:
\begin{align}\label{MainLemmaBPD}
\|\mathbf{A}_j - (\mathbf{BPD})_j\|_2 \leq C_2 \varepsilon, \ \ \  \text{for } j \in [m].
\end{align}
\end{lemma}
